---
title: "Week 02 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exercises

1.  Install the `nycflights13` package and load it into the namespace of your R session. Also, load the `tidyverse` package into the namespace of your R session. Attach the `airlines`, `airports`, `flights`, `weather`, and `planes` datasets

```{r}

```

2.  We'll begin by exploring the `flights` dataset. Determine the following information about the dataset.

-   Structure
-   Dimensions
-   Summary Statistics for each column

```{r}

```

3.  We'll be concentrating on a few key fields. Use the vignette about the `flights` table to look up the definitions of the following fields:

-   `dep_delay` & `arr_delay`
-   `hour` & `minute`
-   `origin`

Copy those definitions as a response here.

```{r}

```

4.  Create a bar chart that demonstrates which origin occurs most frequently. Write a short interpretation of the bar graph.

```{r}

```

5.  What is the distribution of flight departure times by hour? Describe the plot that you created.

    *Hint: Consider different geometries when creating the graph. Histograms are usually a good place to start when considering distributions, but bar plots can be helpful as well.*

```{r}

```

6.  How does the distribution change when we consider the origin of the flight? Build upon your response in question 5. Describe the differences.

    *Hint: Experiment with faceting to find a view that is most helpful.*

```{r}

```

7.  How does departure hour interact with departure delay? Utilize a [boxplot](https://towardsdatascience.com/understanding-boxplots-5e2df7bcbd51). Interpret the graph.

    *Hint: It can be helpful to adjust the portion of the graph that you observe in order to find patterns. Consider using `coord_cartesian` to help with this.*

```{r}

```

8.  Create your own question and answer it by using a graph. You may use any of the datasets loaded in the `nycflights13` package.

-   Indicate what your question is. Consider a question that could be answered best by utilizing facets, multiple geometries, multiple aesthetics, or some combination of those.
-   Create a graph that assists in answering the question.
-   Write a short interpretation of the graph in order to answer your question.

9.  You have now built a variety of graphics using the `ggplot2` system. Describe some of the advantages to the "grammar of graphics" approach to creating graphs.
